---
title: "Differentially Private Generalized Linear Model Release"
author: "James Honaker, Thomas Brawner, and Christian Baehr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dp-glm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Calculate Differentially Private Generalized Linear Model of a set of variables with `glm.release`.

Differential privacy is a rigorous mathematical framework for making statistical information about private datasets available. This is done in such a way that guarantees that information about specific individuals in the dataset does not leak out. Differentially private versions of various statistics are available in this package.  For example, the `glm.release()` function releases coefficient estimates for a set of data after adding a precise about of noise to the objective function to guarantee `epsilon` differential privacy. 

The privacy loss parameter `epsilon` is chosen by the user and represents the degree of privacy preservation guaranteed to each observation in the data when releasing information. Typically this is valued between 0 and 1; as the value gets smaller, the level of privacy protection grows. However, greater privacy protection means more noise must be added to the true mean to achieve the desired amount of privacy. Therefore, as epsilon grows smaller, the privacy protection becomes greater, but the accuracy of the statistical release becomes weaker.

Syntax
------

```{r, eval = FALSE}
private.logit <- glm.release(x = mydata, n = 1000, epsilon = 0.5, formula = y ~ x1 + x2, objective = dp.logit)

private.ols <- glm.release(x = mydata, n = 1000, epsilon = 0.5, formula = y ~ x1 + x2, objective = dp.ols)
```

Inputs
------

`glm.release()` accepts the following arguments to specify the the desired 
level of privacy and characteristics of the variables of interest.

-  `x`: A data frame containing all variables of the model for which the differentially private           coefficient estimates will be calculated.
   
-  `n`: A numeric vector of length one identifying the number of observations
   in `x`. It is important the user specify the number of observations so
   the `glm.release()` can touch the data as little as possible to ensure
   privacy preservation.
   
-  `epsilon`: A numeric vector of length one specifying the epsilon privacy 
   parameter. Typically, this is between zero and one.
   
-  `formula`: An object of class `formula` containing a regression formula using columns names from       `x` to specify variables.
   
-  `objective`: The name of the objective function the user would like to use in the model. Should be:    `dp.ols`, `dp.logit`, `dp.poisson`, or `dp.probit`.

-  `n.boot`: The number of bootstrap samples on which to evaluate the private parameter estimates, each    at `epsilon / n.boot` privacy cost. Default to `NULL`.

-  `intercept`: A logical vector indicating whether an intercept should be added to the model. Default    to `TRUE`.

Examples
------

### Basic Example

Attaching the sample dataset:

```{r, eval = TRUE}
data(PUMS5extract10000, package = "PSIlence")
data <- PUMS5extract10000
```

Calculating differentially private coefficient estimates for a binary dependent variable with a logit model using `glm.release`:

```{r, eval = TRUE}
library(PSIlence)
private.logit <- glm.release(x = data, n = 10000, epsilon = 0.5, formula = married ~ income + age, objective = dp.logit)
private.logit$release
```

To estimate coefficients with a continuous dependent variable instead, use odrinary least squares:

```{r, eval = TRUE}
library(PSIlence)
private.ols <- glm.release(x = data, n = 10000, epsilon = 0.5, formula = income ~ educ + age, objective = dp.ols)
private.ols$release
```

Output Values
------

The output of the `glm.release` function stores fields containing the name of the function call, `glm`,  model type, the objective function of the specified model type, number of observations in the data, the selected epsilon value, the regression formula, whether an intercept was included, and the differentially private release of the coefficients. 