---
title: "Differentially Private Histogram Release"
author: "Megan Fantes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dp-covariance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_knit$set(
        stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 7,
    fig.width = 7
)
```

Differentially Private Histogram with `dpHistogram`
-----------------------------------------

The `dpHistogram` class evaluates a privacy-preserving histogram of a vector of values. The class supports any vector type, meaning that it can handle the R types `numeric`, `integer`, `logical`, and `character`.

<!--
Differential privacy is a rigorous mathematical framework for making statistical information about private datasets available. This is done in such a way that guarantees that information about specific individuals in the dataset does not leak out. Differentially private estimates of various statistics are available in this package.  For example, the `mean.release()` function releases a mean of a variable, while adding a precise about of noise to guarantee `epsilon` differential privacy. 

The privacy loss parameter `epsilon` is chosen by the user and represents the degree of privacy preservation guaranteed to each observation in the data when releasing information. Typically this is valued between 0 and 1; as the value gets smaller, the level of privacy protection grows. However, greater privacy protection means more noise must be added to the true mean to achieve the desired amount of privacy. Therefore, as epsilon grows smaller, the privacy protection becomes greater, but the accuracy of the statistical release becomes weaker.

There is an additional definition of differential privacy known as `epsilon-delta` differential privacy. The additional privacy loss parameter `delta` is chosen by the user and represents how likely the privacy loss for a given dataset is more than epsilon. In other words, `delta` is the chance that additional privacy loss beyond `epsilon` occurs. This value should be very small, on the order of `10^-5`, and never greater than `1/n` (where n is the size of the database). Similar to `epsilon`, as the value of `delta` grows smaller, the level of privacy increases, but the accuracy decreases.

It is noteworthy the user must manually specify the number of observations, variable type, and range of the data of interest; data will only be touched at the point the estimate to be released is actually computed, and only by the function computing that estimate. This structure decreases the possibility of an unauthorized release of information. Additional parameter specification includes the desired `epsilon` value and post-processing steps the user would like to perform on the released statistic. 
-->

Syntax
------

```{r, eval = FALSE}
x1 <- c(3, 12, 20, 42, 33, 65, 70, 54, 33, 45)
data <- data.frame(x1)

example_dpHistogram <- dpHistogram$new(var.type = 'numeric', variable = 'x1', n = 10, epsilon=1, 
                                       rng=c(3,70), n.bins=3, delta=10^-3)

out1 <- example_dpHistogram$release(data)

print(out1)
```

Arguments
---------

In typical usage, there are two methods to the `dpHistogram` class. The `new` method, which creates an object of the class, accepts the following arguments.

* `var.type` \ Character, the type of values in the data frame that will be passed to the mechanism. Should be one of `'numeric'`, `'integer'`, `'logical'`, or `'character'`.

* `variable` \ Character, the name of the variable in the data for which to calculate the histogram.

* `n` \ Integer, the number of observations in the data.

* `epsilon` \ Numeric, the differential privacy parameter $\epsilon$, typically taking values between 0 and 1 and reflecting the privacy cost of the query.

* `accuracy` \ Numeric, the accuracy of the query. Optional, default `NULL`. If `NULL`, the user must specify a value for `epsilon`. If `epsilon` is not `NULL`, this value is ignored and evaluated internally.

* `rng` \ Numeric, a 2-tuple with the lower and upper bounds of the data. Ignored for `'character'` and `'logical'` data types. Optional for `numeric` and `integer` types, default `NULL`. If `NULL` and `bins` also `NULL`, the stability mechanism will be used to calculate the histogram (see ?stabilityMechanism).

* `bins` \ Numeric or character (must match the variable type in `var.type`), a vector of bins for the histogram, as chosen by the user. Optional, default `NULL`. If `NULL`, bins are detected by the histogram statistic. If `NULL` for `'character'` type variables, the detected bins will only be the variables present in the data, and the stability mechanism will be used to calculate the histogram. If `NULL` for `'numeric'` or `'integer'` type variables, the detected bins will be of equal width.

* `n.bins` \ Numeric, the number of bins for the histogram. Ignored for `'character'` and `'logical'` data types. Optional for `numeric` and `integer` types, default `NULL`. If `'NULL'` for `numeric` and `integer` types, `granularity` must be entered.

* `granularity` \ Numeric, the number of items per histogram bin. Ignored for `'character'` and `'logical'` data types. Optional for `numeric` and `integer` types, default `NULL`. If `'NULL'` for `numeric` and `integer` types, `granularity` must be entered.

* `alpha` \ Numeric, the statistical significance level used in evaluating accuracy and privacy parameters. If the bootstrap is employed, `alpha` is also used to trim the release. Default `0.05`.

* `delta` \ Numeric, the differential privacy parameter $\delta$, the likelihood of additional privacy loss beyond $\epsilon$. Typically takes values of $10^{-5}$ or less, should never be less than $1/n$

* `error` \ Numeric, the error term of the statistical significance level. Default is $10^{-9}$.

* `impute.rng` \ Numeric, a 2-tuple giving a range within which missing values of the numeric vector are imputed. Optional, default `NULL`. If `NULL`, missing values are imputed using the range provided in `rng`. Ignored for `character` and `logical` types.

* `impute.bins` \ Character (or numeric for logical variables), a list of bins from which missing (or NA) values of character or logical-type variables will be imputed. Optional, default `NULL`. If `NULL`, missing values are imputed using the histogram bins. Ignored for `numeric` and `integer` types.

* `impute` \ Boolean, if true then the mechanism should replace missing values with known values from the data. If false, the mechanism should leave missing values as `NA`

* `n.boot` \ Integer, the number of bootstrap replications to perform. Optional, default `NULL`. If not `NULL`, the privacy cost `epsilon` is partitioned across `n.boot` replications and the estimates for each are returned.



&nbsp; 

The `release` method accepts a single argument. 

* `data` \ Data frame containing a column corresponding the name specified in `variable`.

Import the PSIlence library and the datasets library, and attach the sample datasets. 

```{r, eval = TRUE}
library(PSIlence)
library(datasets)
data(PUMS5extract10000)
data(SocrataWhiteHouseSalaries)
data(esoph)
```

&nbsp;

Calculate a private histogram of a numeric vector with `dpHistogram`:

```{r}
numeric_histogram <- dpHistogram$new(var.type='numeric', variable='income', n=10000, epsilon=0.1, 
                                     rng=c(0, 750000), n.bins = 5)
numeric_histogram$release(PUMS5extract10000)
print(numeric_histogram$result)
```

In the numeric example above, the histogram statistic detects the buckets for the histogram by creating 5 equal-width buckets for the range of the data. But in the result, we can see that a disproportional number of the data points are in the first bucket. This makes sense, as this is income data, and a large portion of people have an income less than \$150,000 dollars (the range of the first bucket). If we have prior knowledge of the data, we can enter specific buckets for the histogram in the `bins` parameter. For example, let's say we want a histogram of incomes less than or equal to \$50,000; \$50,000 - \$100,000; \$100,000 - \$200,000; \$200,000 - \$300,000; \$300,000 - \$400,000; and > \$400,000:

```{r}
# When entering bins, enter the minimum of the range, the upper bound of each bin, and the maximum of the range
incomeBins <- c(0,50000, 100000, 200000, 300000, 400000, 750000)

# So in this example, we have 6 bins:
# 1. (min, 500000]
# 2. (50000, 100000]
# 3. (100000, 200000]
# 4. (200000, 300000]
# 5. (300000, 400000]
# 6. (400000, max]
# where min = the minimum of the dataset = 0
# and max = the maximum of the dataset = 750000

numeric_histogram <- dpHistogram$new(var.type='numeric', variable='income', n=10000, epsilon=0.1, 
                                     rng=c(0,750000), n.bins=6, bins=incomeBins)
numeric_histogram$release(PUMS5extract10000)
print(numeric_histogram$result)
```

&nbsp;

To calculate the histogram of a logical variable, input a logical vector into `variable` and update `var.type` to 'logical'. Also add a boolean flag for `impute` to indicate if you would like empty values to be imputed from the existing data. (If `impute = TRUE`, then there will only me two bins in the histogram: 0 and 1. If `impute = FALSE`, then there will be three bins in the histogram: 0, 1, and NA.)

```{r}
logical_histogram <- dpHistogram$new(var.type='logical', variable='married', n=10000, epsilon=0.1, impute = TRUE)
logical_histogram$release(PUMS5extract10000)
print(logical_histogram$result)
```

In the example above, we set the `impute` parameter equal to `TRUE`, meaning that any `NA` (empty) values would be randomly assigned 0 or 1 so no values in the data set are empty, and there is no `NA` bucket in the histogram. By default `impute = FALSE`, so if we do not specify a value for `impute`, then there will be an `NA` bucket in the histogram release:

```{r}
logical_histogram <- dpHistogram$new(var.type='logical', variable='married', n=10000, epsilon=0.1)
logical_histogram$release(PUMS5extract10000)
print(logical_histogram$result)
```

For logical variables, it is never necessary to enter a vector of bins, you just need to specify whether to impute empty values or not.

&nbsp;

To calculate the histogram of a categorical (or 'character') variable, input a character variable name into `variable` and update `var.type` to 'character'.

```{r}
character_histogram <- dpHistogram(var.type='character', variable='tobgp', n=88, epsilon=1, delta=10^-3)
character_histogram$release(esoph)
print(character_histogram)
```

Notice that in the character histogram example above, the sample size is very small, only 88 observations, so the `epsilon` and `delta` parameters must be much larger to accomodate the small data size. A large `epsilon` and `delta` indicate less privacy and more accuracy than the small values used in earlier examples with larger sample sizes. However, less privacy indicates more accuracy, which is necessary with small sample sizes that are sensitive to the differentially private noise added during calculation.

Similar to numeric variables, you can enter a specific vector of bins for the histogram for a character variable if you have prior knowledge of the data. In this case, the resulting histogram will have each bin entered in the `bins` vector, even if the bin would be empty in the underlying data:

```{r}
character_histogram <- dpHistogram(var.type='character', variable="Employee.Status", n=469, epsilon=0.5, delta=10^-5,
                                   bins=c("Detailee","Employee","Employee (part-time)"))
character_histogram$release(SocrataWhiteHouseEmployeeSalaries)
print(character_histogram)
```

&nbsp;

NEED TO UPDATE THIS ONCE PULL REQUEST IS MERGED --->

Values
------

The `release` method makes a call to the mechanism, which generates a list of statistical summaries available on the `result` field.

* `result` List, contains the accuracy guarantee, privacy cost, and private release. Other elements reflecting variable post-processing of the release.

&nbsp; 

The list in the `result` attribute has the following values.

* `release` \ Differentially private estimate of the mean. If the bootstrap mechanism is used, one estimate for each bootstrap replication is provided (i.e., vector of length `n.boot`).
* `accuracy` \ The accuracy guarantee of the release given `epsilon`.
* `epsilon` \ The privacy cost required to guarantee `accuracy`.
* `interval` \ Confidence interval of the private estimate given `accuracy`.
* `std.dev` \ The standard deviation of the vector. Only available for logical vectors.
* `median` \ The median of the vector. Only available for logical vectors.
* `histogram` \ The histogram of the vector. Only available for logical vectors.
* `std.error` \ Estimates of the standard error of the mean. Only available when the bootstrap mechanism is used.