---
title: "Differentially Private Mean Release"
author: "James Honaker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dp-mean}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Calculate Differentially Private Mean of a Vector with `mean.release`.

Differential privacy is a rigorous mathematical framework for making statistical information about private datasets available. This is done in such a way that guarantees that information about specific individuals in the dataset does not leak out. The `mean.release()` function releases the differentially private mean of a variable while guaranteeing `epsilon` differential privacy. Depending on the value of `epsilon`, a degree of noise is added to the true mean to make it differentially private.

`epsilon` is chosen by the user and represents the degree of privacy given to each observation in the data when releasing information about the data. The epsilon value should be between zero and one, and as the value gets smaller, the level of privacy protection grows. However, greater privacy protection means more noise must be added to the true mean to achieve the desired amount of privacy. Therefore, as epsilon grows smaller, the privacy protection becomes greater, but the accuracy guarantee statistical release becomes weaker.

Syntax
------

```{r, eval = FALSE}
mydata <- c(3, 12, 20, 42, 33, 65, 70, 54, 33, 45)
private.mean <- mean.release(x = mydata, var.type = "numeric", epsilon = 0.1, n = 10, rng = c(3, 70))

mydata2 <- c(TRUE, FALSE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE)
private.mean2 <- mean.release(x = mydata2, var.type = "logical", epsilon = 0.5, n = 8, rng = c(0, 1))
```

Inputs
------

`mean.release()` accepts the following arguments to specify the the desired 
level of privacy and characteristics of the variable of interest.

-  `x`: The vector of data for which the differentially private mean will
   be calculated. This data can be of type numeric, integer, or
   logical.
   
-  `var.type`: A character vector of length one identifying the data type
   of `x`. This field should contain either 'numeric', 'integer', or
   'logical'.
   
-  `epsilon`: A numeric vector of length one specifying the epsilon privacy 
   parameter. This should be between zero and one.
   
-  `n`: A numeric vector of length one identifying the number of observations
   in `x`. It is important the user specify the number of observations so
   the `mean.release()` can touch the data as little as possible to ensure
   privacy preservation.
   
-  `rng`: A numeric vector of length two identifying the range of `x`. If this
   vector is greater than length two, the maximum and minimum values of this vector
   will be used.

Examples
------

### Basic Example







Attaching the sample dataset:

```{r, eval = TRUE}
data(tobin)
```

Estimating linear regression using `tobit`:

```{r, eval = TRUE}
z.out <- zelig(durable ~ age + quant, model = "tobit", data = tobin)
```

Summarize estimated paramters:

```{r, eval = TRUE}
summary(z.out)
```

Setting values for the explanatory variables to their sample averages:

```{r, eval = TRUE}
x.out <- setx(z.out)
```

Simulating quantities of interest from the posterior distribution given `x.out`.

```{r, eval = TRUE}
s.out1 <- sim(z.out, x = x.out)
```

```{r, eval = TRUE}
summary(s.out1)
```








- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
