---
title: "Differentially Private Mean Release"
author: "James Honaker, Thomas Brawner, and Christian Baehr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dp-mean}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_knit$set(
        stop_on_error = 2L
)
knitr::opts_chunk$set(
    fig.height = 7,
    fig.width = 7
)
```

Differentially Private Mean with `dpMean`
-----------------------------------------

The `dpMean` class evaluates a privacy-preserving mean of a vector of values. The class supports any vector type that can be represented numerically, meaning that it can handle the R types `numeric`, `integer`, and `logical`.

<!--
Differential privacy is a rigorous mathematical framework for making statistical information about private datasets available. This is done in such a way that guarantees that information about specific individuals in the dataset does not leak out. Differentially private estimates of various statistics are available in this package.  For example, the `mean.release()` function releases a mean of a variable, while adding a precise about of noise to guarantee `epsilon` differential privacy.

The privacy loss parameter `epsilon` is chosen by the user and represents the degree of privacy preservation guaranteed to each observation in the data when releasing information. Typically this is valued between 0 and 1; as the value gets smaller, the level of privacy protection grows. However, greater privacy protection means more noise must be added to the true mean to achieve the desired amount of privacy. Therefore, as epsilon grows smaller, the privacy protection becomes greater, but the accuracy of the statistical release becomes weaker.

It is noteworthy the user must manually specify the number of observations, variable type, and range of the data of interest; data will only be touched at the point the estimate to be released is actually computed, and only by the function computing that estimate. This structure decreases the possibility of an unauthorized release of information. Additional parameter specification includes the desired `epsilon` value and post-processing steps the user would like to perform on the released statistic.
-->

Syntax
------

```{r, eval = FALSE}
library(psilence)
data(PUMS5extract10000)

dpMeanExample <- dpMean$new(mechanism='mechanismLaplace', varType='numeric',
                      variable='income', epsilon=0.1, n=nrow(PUMS5extract10000), rng=c(0, 250000))
dpMeanExample$release(PUMS5extract10000)
print(dpMeanExample$result)

dpMeanExample2 <- dpMean$new(mechanism='mechanismLaplace', varType='logical',
                       variable='sex', epsilon=0.1, n=nrow(PUMS5extract10000), rng=c(0, 1))
dpMeanExample2$release(PUMS5extract10000)
print(dpMeanExample2$result)

dpMeanExample3 <- dpMean$new(mechanism='mechanismSnapping', varType='numeric',
                      variable='income', epsilon=0.1, n=nrow(PUMS5extract10000), rng=c(0, 250000))
dpMeanExample3$release(PUMS5extract10000)
print(dpMeanExample3$result)

dpMeanExample4 <- dpMean$new(mechanism='mechanismSnapping', varType='logical',
                       variable='sex', epsilon=0.1, n=nrow(PUMS5extract10000), rng=c(0, 1))
dpMeanExample4$release(PUMS5extract10000)
print(dpMeanExample4$result)
```

Arguments
---------

In typical usage, there are two methods to the `dpMean` class. The `new` method, which creates an object of the class, accepts the following arguments.

* `mechanism` \ Character, the class name of the mechanism used to perturb the true estimate, one of `'mechanismLaplace'`, `'mechanismSnapping'`, or `'mechanismBootstrap'`.

* `varType` \ Character, the type of values in the data frame that will be passed to the mechanism. Should be one of `'numeric'`, `'integer'`, or `'logical'`.

* `n` \ Integer, the number of observations in the vector.

* `rng` \ Numeric, a 2-tuple giving an a priori estimate of the lower and upper bounds of the vector.

* `epsilon` \ Numeric, the differential privacy parameter $\epsilon$, typically taking values between 0 and 1 and reflecting the privacy cost of the query. Optional, default `NULL`. If `NULL`, the user must specify a value for `accuracy`.

* `accuracy` \ Numeric, the accuracy of the query. Optional, default `NULL`. If `NULL`, the user must specify a value for `epsilon`. If `epsilon` is not `NULL`, this value is ignored and evaluated internally.

* `imputeRng` \ Numeric, a 2-tuple giving a range within which missing values of the vector are imputed. Optional, default `NULL`. If `NULL`, missing values are imputed using the range provided in `rng`.

* `nBoot` \ Integer, the number of bootstrap replications to perform. Optional, default `NULL`. If not `NULL`, the privacy cost `epsilon` is partitioned across `nBoot` replications and the estimates for each are returned.

* `alpha` \ Numeric, the statistical significance level used in evaluating accuracy and privacy parameters. If the bootstrap is employed, `alpha` is also used to trim the release. Default `0.05`.

* `gamma` \ Numeric, used to provide high probability (1-gamma) guarantee that clamping bound does not bind. Optional, used only for Snapping Mechanism. Default `0.05`.

&nbsp;

The `release` method accepts a single argument.

* `x` \ Data frame containing numeric columns corresponding the names specified in `formula`.


Examples
------

Attach the sample dataset.

```{r, eval = TRUE}
library(PSIlence)
data(PUMS5extract10000)
```

&nbsp;

Calculate a private mean of a numeric vector with `dpMean` using the Laplace mechanism:

```{r}
numericMean <- dpMean$new(mechanism='mechanismLaplace', varType='numeric',
                           variable='income', n=10000, epsilon=0.1, rng=c(0, 750000))
numericMean$release(PUMS5extract10000)
print(numericMean$result)
```

&nbsp;

To calculate the mean of a logical vector instead, input a logical vector into `x` and update `varType` and `rng` appropriately:

```{r}
logicalMean <- dpMean$new(mechanism='mechanismLaplace', varType='logical',
                           variable='married', n=10000, epsilon=0.1, rng=c(0, 1))
logicalMean$release(PUMS5extract10000)
print(logicalMean$result)
```

Values
------

The `release` method makes a call to the mechanism, which generates a list of statistical summaries available on the `result` field.

* `result` List, contains the accuracy guarantee, privacy cost, and private release. Other elements reflecting variable post-processing of the release.

&nbsp;

The list in the `result` attribute has the following values.

* `release` \ Differentially private estimate of the mean. If the bootstrap mechanism is used, one estimate for each bootstrap replication is provided (i.e., vector of length `nBoot`).
* `accuracy` \ The accuracy guarantee of the release given `epsilon`.
* `epsilon` \ The privacy cost required to guarantee `accuracy`.
* `interval` \ Confidence interval of the private estimate given `accuracy`.
* `std.dev` \ The standard deviation of the vector. Only available for logical vectors.
* `median` \ The median of the vector. Only available for logical vectors.
* `histogram` \ The histogram of the vector. Only available for logical vectors.
* `std.error` \ Estimates of the standard error of the mean. Only available when the bootstrap mechanism is used.
