---
title: "Differentially Private Mean Release"
author: "James Honaker, Thomas Brawner, and Christian Baehr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dp-mean}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Calculate Differentially Private Mean of a Vector with `mean.release`.

Differential privacy is a rigorous mathematical framework for making statistical information about private datasets available. This is done in such a way that guarantees that information about specific individuals in the dataset does not leak out. Differentially private versions of various statistics are available in this package.  For example, the `mean.release()` function releases a mean of a variable, while adding a precise about of noise to guarantee `epsilon` differential privacy. 

The privacy loss parameter `epsilon` is chosen by the user and represents the degree of privacy preservation guaranteed to each observation in the data when releasing information. Typically this is valued between 0 and 1; as the value gets smaller, the level of privacy protection grows. However, greater privacy protection means more noise must be added to the true mean to achieve the desired amount of privacy. Therefore, as epsilon grows smaller, the privacy protection becomes greater, but the accuracy of the statistical release becomes weaker.

Syntax
------

```{r, eval = FALSE}
mydata <- c(3, 12, 20, 42, 33, 65, 70, 54, 33, 45)
private.mean <- mean.release(x = mydata, var.type = "numeric", epsilon = 0.1, n = 10, rng = c(3, 70))

mydata2 <- c(TRUE, FALSE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE)
private.mean2 <- mean.release(x = mydata2, var.type = "logical", epsilon = 0.5, n = 8, rng = c(0, 1))
```

Inputs
------

`mean.release()` accepts the following arguments to specify the the desired 
level of privacy and characteristics of the variable of interest.

-  `x`: The vector of data for which the differentially private mean will
   be calculated. This data can be of type numeric, integer, or
   logical.
   
-  `var.type`: A character vector of length one identifying the data type
   of `x`. This field should contain either 'numeric', 'integer', or
   'logical'.
   
-  `epsilon`: A numeric vector of length one specifying the epsilon privacy 
   parameter. Typically, this is between zero and one.
   
-  `n`: A numeric vector of length one identifying the number of observations
   in `x`. It is important the user specify the number of observations so
   the `mean.release()` can touch the data as little as possible to ensure
   privacy preservation.
   
-  `rng`: A numeric vector of length two identifying the range of `x`. If this
   vector is greater than length two, the maximum and minimum values of this vector
   will be used.

Examples
------

### Basic Example

Attaching the sample dataset:

```{r, eval = TRUE}
data(PUMS5extract10000, package = "PSIlence")
data <- PUMS5extract10000
```

Calculating the differentially private mean of a numeric vector using `mean.release`:

```{r, eval = TRUE}
library(PSIlence)
private.numeric <- mean.release(x = data$income, var.type = "numeric", n = 10000, 
                             epsilon = 0.01, rng = c(-10000, 713000))
private.numeric$release
```

To calculate the mean of a logical vector instead, input a logical vector into `x` and update `var.type` and `rng`:

```{r, eval = TRUE}
library(PSIlence)
private.logical <- mean.release(x = data$married, var.type = "logical", n = 10000, 
                             epsilon = 0.01, rng = c(0, 1))
private.logical$release
```

Output Values
------

The output of the `mean.release` function stores fields containing the name of the statistic, variable type, number of observations, the sensitivity of the data, the selected epsilon value, the differentially private release of the mean, the accuracy guarantee of the mean release, and the bounds of the confidence interval for the mean release. 
